Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-19 cav2017/cav2017.tex
restructuring

The closed-loop state-space system has to be unwound up to a given 
depth $k_{bound}$ to ensure that $\phi_{safety}$ holds. 
Here, $k_{bound}$ represents the number of iterations 
to sufficiently unwind the closed-loop state-space system 
so that the variation rate (derivative) of the states is negligible, 
ensuring that the boundaries are not be violated for any $k>k_{bound}$. 
\commentib{The k-bound estimation must be explained.}

For any $k\leq k_{bound}$, the states must be computed considering 
the FWL effects in all operations related to the control action~\eqref{eq:controlaction} 
and disregarding it in the state equation~\eqref{eq:plant}. 
Thus, the states are computed using the following equations:
%
\begin{equation}
\label{eq:ssunroll}
\left\lbrace
\begin{array}{l}
u_{k}=\mathcal{F}_{\langle I,F \rangle}(r_{k}-\mathcal{F}_{\langle I,F \rangle}(\mathcal{F}_{\langle I,F \rangle}(K)\cdot\mathcal{F}_{\langle I,F \rangle}(x_{k})))\\
x_{k+1}=Ax_{k}+Bu_{k}
\end{array}\right.
\end{equation}

@@

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Synthesising the dynamics}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
\begin{enumerate}
\item We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{A}_o,\mat{B}_o,\mat{C}_o$ by rounding down the coefficients $*_{oij}=\mathcal{F}\langle I,F,E\rangle (*_{ij}) : *=\{a,b,c\}$
\item Calculate $\mat{T}$ and $\mat{W}_o^{-1}\mat{W}_{of}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}\eqref{eq:wnof}\eqref{eq:wof}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=(\mat{T}\mat{A}\mat{T}^-1)_{n*}$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\item Produce an initial abstraction which corresponds to an unconstrained solution space.
\end{enumerate}
\item Next we synthesise a candidate controller and observer. Let $\mathcal{J}\langle I,F,E \rangle (P,\mat{T})$ be a program describing Jury's method.
\begin{enumerate}
\item Select a controller $\tilde{\mat{K}} \in \mathbb{R}\langle I,F,E\rangle^n$. This is achieved by solving a SAT formula $\mathcal{J}_K=\mathcal{J}\langle I,F,E \rangle (P_{a-k},\mat{T})$ that satisfies Jury's criteria for $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \mat{K}=\tilde{\mat{K}}\mat{T}$.
\item Select the observer error characteristic polynomial $\tilde{\mat{L}} \in \mathbb{R}\langle I,F,E\rangle^n$. 
This is achieved by solving a SAT formula $\mathcal{J}_L=\mathcal{J}\langle I,F,E \rangle (P_p,\mat{W}_o^{-1}\mat{W}_{of})$ 
that satisfies Jury's criteria for 
$P_p=z^n+\sum_{i=1}^n p_i z^{n-i} : p_i \in \vec{P}= (\mat{W}_o^{-1}\mat{W}_{of})^-1 \tilde{\mat{L}} 
+ [\begin{array}{cccc}a_1&a_2&\hdots&a_n\end{array}]^T$.
\item If there is no candidate solution we return UNSAT and exit the loop.
\end{enumerate}
\item Once we have a candidate solution, we perform a safety verification using abstract acceleration.
\begin{enumerate}
\item Calculate the abstract matrices $\mathcal{A}_t, \mathcal{B}_{tp},\mathcal{B}_{tn}$ for the candidate closed loop solution.
\item Evaluate $\hat{X}^\# \models \Psi$. If the verification holds we have a solution, and exit the loop.
\end{enumerate} 
\item If the candidate solution is not valid we refine the abstraction used by the synthesiser
\begin{enumerate}
\item Find the constraints that invalidate the property. These are the rows of $\mat{E} \in \Psi$ that intersect/segment $\hat{X}^\#$. Transforming these rows into the eigenspace of $\mat{A}_t$, and applying backward acceleration to obtain a set of counterexamples for the eigenvalues, which we define as $\Psi(\Lambda)$. This is a constraint in the spectrum (\emph{ie} frequency response transfer function) of the closed loop dynamics. 
\item Transform the soft constraints on the spectrum into soft constraints on the coefficients of the characteristic polynomial. 
$z^n+\sum_{i=1}^n[a_i-k_i]z^{n-i}=\prod_{i=1}^n (z-[\lambda_i]) : |[\lambda_i]|<1 \wedge [\lambda_i] \models \Psi_{\Lambda}$.
\item Pass the refined abstraction with the new contraints to the synthesiser, as well as a counterexample plant coefficients ($P_a$) and repeat the loop again. We note that the Abstraction refinement is a soft constraint on the proposed model which can increase or decrease the parameter space at any time, whereas the concrete counterexample passed to the synthesiser is monotonically decreasing. The first one improves speed of the tool whereas the second ensures convergence.

@@

 \end{tikzpicture}
}}
\caption{The CEGIS loop with precision and sampling rate incrementation}
\label{fig:CEGIS}
\end{figure}

>>>>

The states must be computed considering 
the FWL effects in all operations related to the control action~\eqref{eq:controlaction} 
and disregarding it in the state equation~\eqref{eq:plant}. 
Thus, the states are computed using the following equations:
%
\begin{equation}
\label{eq:ssunroll}
\left\lbrace
\begin{array}{l}
u_{k}=\mathcal{F}_{\langle I,F \rangle}(r_{k}-\mathcal{F}_{\langle I,F \rangle}(\mathcal{F}_{\langle I,F \rangle}(K)\cdot\mathcal{F}_{\langle I,F \rangle}(x_{k})))\\
x_{k+1}=Ax_{k}+Bu_{k}
\end{array}\right.
\end{equation}

@@

\subsubsection{Computing the set of reachable states}
\todo{explain the use of abstract acceleration}

@@

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsubsection{Controller synthesis}\label{sec:cof_synth}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
Now that we have explained the verification process, we may proceed to describe the synthesis procedure, which is illustrated in figure \ref{fig:CEGARIS}.
\begin{enumerate}
\item We start by preprocessing the dynamics. This can be computationally intensive and we don't want it to be part of the CEGIS loop.
\begin{enumerate}
\item Calculate $\mat{A}_o,\mat{B}_o,\mat{C}_o$ by rounding down the coefficients $*_{oij}=\mathcal{F}\langle I,F,E\rangle (*_{ij}) : *=\{a,b,c\}$
\item Calculate $\mat{T}$ and $\mat{W}_o^{-1}\mat{W}_{of}$ using equations \eqref{eq:rncf}\eqref{eq:rcf}\eqref{eq:to_cf}\eqref{eq:wnof}\eqref{eq:wof}
\item Extract the Plant's characteristic polynomial coefficients from the last row of $P_a=(\mat{T}\mat{A}\mat{T}^-1)_{n*}$.
\item Calculate the noise set $N$ from the quantizer resolutions and estimated round-off errors ($\frac{q_1}{2},\frac{q_2}{2},q_3$)
\item Produce an initial abstraction which corresponds to an unconstrained solution space.
\end{enumerate}
\item Next we synthesise a candidate controller and observer. Let $\mathcal{J}\langle I,F,E \rangle (P,\mat{T})$ be a program describing Jury's method.
\begin{enumerate}
\item Select a controller $\tilde{\mat{K}} \in \mathbb{R}\langle I,F,E\rangle^n$. This is achieved by solving a SAT formula $\mathcal{J}_K=\mathcal{J}\langle I,F,E \rangle (P_{a-k},\mat{T})$ that satisfies Jury's criteria for $P_{a-k}=z^n+\sum_{i=1}^n (a_i-k_i) z^{n-i} : k_i \in \mat{K} \wedge \mat{K}=\tilde{\mat{K}}\mat{T}$.
\item Select the observer error characteristic polynomial $\tilde{\mat{L}} \in \mathbb{R}\langle I,F,E\rangle^n$. 
This is achieved by solving a SAT formula $\mathcal{J}_L=\mathcal{J}\langle I,F,E \rangle (P_p,\mat{W}_o^{-1}\mat{W}_{of})$ 
that satisfies Jury's criteria for 
$P_p=z^n+\sum_{i=1}^n p_i z^{n-i} : p_i \in \vec{P}= (\mat{W}_o^{-1}\mat{W}_{of})^-1 \tilde{\mat{L}} 
+ [\begin{array}{cccc}a_1&a_2&\hdots&a_n\end{array}]^T$.
\item If there is no candidate solution we return UNSAT and exit the loop.
\end{enumerate}
\item Once we have a candidate solution, we perform a safety verification using abstract acceleration.
\begin{enumerate}
\item Calculate the abstract matrices $\mathcal{A}_t, \mathcal{B}_{tp},\mathcal{B}_{tn}$ for the candidate closed loop solution.
\item Evaluate $\hat{X}^\# \models \Psi$. If the verification holds we have a solution, and exit the loop.
\end{enumerate} 
\item If the candidate solution is not valid we refine the abstraction used by the synthesiser
\begin{enumerate}
\item Find the constraints that invalidate the property. These are the rows of $\mat{E} \in \Psi$ that intersect/segment $\hat{X}^\#$. Transforming these rows into the eigenspace of $\mat{A}_t$, and applying backward acceleration to obtain a set of counterexamples for the eigenvalues, which we define as $\Psi(\Lambda)$. This is a constraint in the spectrum (\emph{ie} frequency response transfer function) of the closed loop dynamics. 
\item Transform the soft constraints on the spectrum into soft constraints on the coefficients of the characteristic polynomial. 
$z^n+\sum_{i=1}^n[a_i-k_i]z^{n-i}=\prod_{i=1}^n (z-[\lambda_i]) : |[\lambda_i]|<1 \wedge [\lambda_i] \models \Psi_{\Lambda}$.
\item Pass the refined abstraction with the new contraints to the synthesiser, as well as a counterexample plant coefficients ($P_a$) and repeat the loop again. We note that the Abstraction refinement is a soft constraint on the proposed model which can increase or decrease the parameter space at any time, whereas the concrete counterexample passed to the synthesiser is monotonically decreasing. The first one improves speed of the tool whereas the second ensures convergence.

@@

 \end{tikzpicture}
}}
\caption{The CEGIS loop with precision and sampling rate incrementation}
\label{fig:CEGIS-precision-increment}
\end{figure}

\subsubsection{Computing the reachable states}

In order to compute the set of reachable states we find 
a completeness threshold $k_{bound}$.
Essentially, $k_{bound}$ represents the number of iterations 
to sufficiently unwind the closed-loop state-space system 
so that the variation rate (derivative) of the states is negligible, 
ensuring that the boundaries are not be violated for any $k>k_{bound}$. 
The closed-loop state-space system has to be unwound up to a given 
depth $k_{bound}$ to ensure that $\phi_{safety}$ holds.

\todo{discuss the completeness threshold computation}

\subsubsection{Controller synthesis}

\todo{Describe each step in Fig~\ref{fig:CEGIS-precision-increment}.}
